Safe and Robust Formal Synthesis of Digital Controllers for Continuous Plants

### 2018-02-26 automatica/paper.tex
removed response time and transient analysis terms
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

\begin{keyword}                           % Five to ten keywords,  
Digital Control; A/D converters; Control System Synthesis; Transient Analysis; Safety Analysis; Quantization Errors; Sampling.               % chosen from the IFAC 
\end{keyword}                             % keyword list or with the 
                                          % help of the Automatica 
                                          % keyword wizard

@@

Our work addresses challenging aspects of the control synthesis problem.  We
perform digital control synthesis over a hybrid model, where the plant
exhibits continuous behavior whereas the controller operates in discrete
time and over a quantized domain.  In particular, our model evaluates the
effects of the quantizers (A/D and D/A converters) and of time
discretization, as well as representation errors introduced by in modelling
artifacts such as an observer~\cite{astrom1997computer} in a finite
precision domain, while reasoning about high level properties such as
stability, robustness, safety, and response time, which define the behaviour
of the system.  In particular, safety requirements are frequently overlooked
in conventional feedback control synthesis, but play an important role in
systems engineering.

@@

  % Additions from CAV:
  % continuous time
  % response time
  % floating point ?
\item We automatically generate {\em correct-by-construction} digital
  controllers using an inductive synthesis approach.  Our application of program
  synthesis is non-trivial and addresses challenges specific to control
  systems, such as quantizers and FWL.  %% In particular, we
  %% have found that a two-stage verification engine that continuously refines
  %% the precision of the fixed-point representation of the plant yields a
  %% speed-up of two orders of magnitude over a conventional one-stage
  %% verification engine.
%
  %\item  We compute
  The computed state-feedback controllers 
  guarantee a given safety property alongside stability and response time
  specifications.  Existing methods for controller synthesis rely on transfer
  function representations, which are inadequate to prove safety requirements.
%
\item We provide two novel algorithms: the first, na\"ive one, relies on
  an unfolding of the dynamics up to a completeness threshold, while the
  second one is abstraction-based and leverages abstraction refinement and
  acceleration to improve scalability while retaining soundness.  %% A third
  %% algorithm evaluates systems using transfer functions where no state-space
  %% description is given.
  Both approaches provide sound synthesis of
  feedback controllers and consider the various sources of imprecision in
  the implementation of the control algorithm and in the modeling of the
  plant.
%
\item We develop a model for different sources of quantization error and
  their effect on reachability properties.  We give bounds that ensure the
  safety of our controllers in a hybrid continuous-digital domain.
\end{enumerate}

@@

%-------------------------------
\subsection{Formal specification of transient response} 
\label{ssec:transientspecification}
%-------------------------------

In a SISO system, there are 2 values that typically define the response-time of the system (using a step response).
\begin{enumerate}
\item The \emph{rise time} is defined as the time it takes the output to reach a percentage (eg $90\%$) of the target value given a step chnage in the input.
\item The \emph{settling time} is defined as the time it takes the output to stabilize within a margin of the target value (eg $10\%$).
\end{enumerate} 
In the case of first and second order systems (with ), the decay is governed by the equations 
\begin{align*}
&y_{decay}=
&\left\{
\begin{array}{cc}
e^{-|\omega| t}& 1^{st} order\\
e^{\zeta |\omega| t}\left(\frac{\sin\left(|\omega| t\sqrt{1-\zeta^2}+\sin^{-1}\sqrt{1-\zeta^2}\right)}{\sqrt{1-\zeta^2}} \right)&\zeta<1\\
e^{-|\omega| t}-|\omega| t e^{-|\omega| t}&\zeta=1\\
\frac{|\omega| e^{-|\omega_2| t}-|\omega_2| e^{-|\omega| t}}{|\omega|-|\omega_2|}&\zeta>1\end{array}\right.
\end{align*}
where $|\omega|$ and $|\omega_2|$ are the magnitudes of the eigenvalues (roots) of the system and  $\zeta=-\cos(\theta)$ where $\theta$ is the polar angle of the conjugate pair if it exists ($\zeta \geq 1$ if it doesn't).
These equations become increasingly complicated for higher order systems. However, we note their strong dependency on the magnitude of the eigenvalues, and use this property to define an upperbound on the decay function.
All of the above equations will decay faster than $e^{-|\omega| t}$ given $|\omega| < |\omega_2|$, hence we may use this function as our upper bound.
\begin{theorem}
Given a set of eigenvalues for a discrete time model of an $n^{th}$ order system $\lambda_i =e^{\omega_i}: i \in [1\ n]$, the settling time of the system is upperbounded by
\begin{equation}
t_s<-log_{|\overline{\lambda}|}({p_s}) : |\overline{\lambda}| =\max(|\lambda_i|) \wedge |\overline{\lambda}|<1,\ \forall i \in [1\ n]
\label{eq:set_time}
\end{equation}
where $0\leq p_s \leq 1$ is the desired percentage of the target within to remain after $t_s$.
\end{theorem} 
\begin{proof}
Given a closed loop single output LTI system of order $n$ with dynamics
\begin{equation}
\left\{\begin{array}{c}\vec{x}_k=\mat{A}\vec{x}_{k-1}\\ \vec{y}_k=\mat{C}\vec{x}_k\end{array}\right. \Rightarrow \left\{ \begin{array}{c}\vec{x}_k'=\mat{J}\vec{x}_{k-1}'\\ \vec{y}_k=\mat{C}\mat{S}\vec{x}_k'\end{array}\right. : \left\{ \begin{array}{c}\mat{A}=\mat{S}\mat{J}\mat{S}^{-1} \\ \vec{x}_k'=\mat{S}^{-1}\vec{x}_k\end{array}\right.
\label{eq:eigendynamics}
\end{equation}
%
the output decay is a linear combination of the decay of eigenvalues or
eigenvalue pairs.  
Let $\mat{C}'\in \mathbb{R}^{1 \times n}=\mat{C}\mat{S}$. Since $\mat{C}'$ is constant, the output $y_k$ is a linear combination of the states $\vec{x}'_k$. We accelerate equation \eqref{eq:eigendynamics} to obtain
\begin{align}
%\vec{x}_k'&=\mat{J}^k\vec{x}_0' \\
y_k&=\mat{C}'\mat{J}^k\vec{x}_0 = \sum_{i=0}^n \left( c_i \lambda_i^k + c_{il}\binom{k}{l} \lambda_i^{k-l}\right)\\
|y_k|&\leq \sum_{i=0}^n \left|c_i + c_{il}\binom{k}{l} \lambda_i^{-l}\right||\lambda_i^k|
\end{align}
where the second  part of the equation corresponds to the upper diagonal terms of the Jordan blocks. We also have $|\lambda_i^k|\leq |\overline{\omega}^k|$,hence
\begin{align}
|y_k|&\leq |\overline{\omega}^k|\sum_{i=0}^n \left|c_i + c_{il}\binom{k}{l} \lambda_i^{-l}\right|
\end{align}
Since we are synthesising the eigenvalues, rather than solve this formula for complex Jordan Shapes, we will restrict our synthesiser to provide solutions with distinct eigenvalues (which is by far the most common case). Therefore the right hand terms dissappear and we end up with
\begin{align}
|y_k|&\leq |\overline{\omega}^k|\sum_{i=0}^n |c_i|
\end{align}

Hence the overall decay of the system is always smaller
than $p=e^{-|\overline{\omega}| t}$ where $0\leq p\leq 1$ is the remaining
percentage of the decay expressed.  reversing we have
$t=-\frac{ln(p)}{|\overline{\omega}|}=-\frac{ln(p)}{ln(|\overline{\lambda}|)}$. 
replacing for $p_s$ , $t_s$, and $\overline{\lambda}$ we get equation
\eqref{eq:set_time}
%
\qed
\end{proof}
%
We may now use this limit to express a constraint in terms of the
coefficients for our synthesis process.  Let
%
\begin{equation}
P=\prod_{i=0}^n (s-\lambda_i)=\sum_{j=0}^n c_js^j.
\end{equation}
%
We use Jury's criterion to limit $|\lambda_i| \leq 1$.  Let us now define a
transformation $\lambda_i'=\frac{\lambda_i}{\overline{\lambda}}$.  We want
to ensure $\lambda_i' \leq 1$ which can be done by applying Jury's criterion
to $P'=\prod_{i=0}^n (s-\lambda_i')=\sum_{j=0}^n c_j's^j$ where $c_j'$ are
unknown.  However, if we replace for $\lambda_i$, we obtain
%
\begin{equation}
P'=\prod_{i=0}^n (s-\lambda_i')=\prod_{i=0}^n \left(s-\frac{\lambda_i}{\overline{\lambda}}\right)=\sum_{j=0}^n \frac{c_j}{\overline{\lambda}^{n-j}}s^j
\end{equation}
%
where all members are known, we may now synthesise coefficients that ensure
the eigenvalues are smaller than a desired magnitue, thus ensuring that the
settling time is too.

>>>>

\begin{keyword}                           % Five to ten keywords,  
Digital Control; A/D converters; Control System Synthesis; Safety Analysis; Quantization Errors; Sampling.               % chosen from the IFAC 
\end{keyword}                             % keyword list or with the 
                                          % help of the Automatica 
                                          % keyword wizard

@@

Our work addresses challenging aspects of the control synthesis problem.  We
perform digital control synthesis over a hybrid model, where the plant
exhibits continuous behavior whereas the controller operates in discrete
time and over a quantized domain.  In particular, our model evaluates the
effects of the quantizers (A/D and D/A converters) and of time
discretization, as well as representation errors introduced by in modelling
artifacts such as an observer~\cite{astrom1997computer} in a finite
precision domain, while reasoning about high level properties such as
stability, robustness, and safety, which define the behaviour
of the system.  In particular, safety requirements are frequently overlooked
in conventional feedback control synthesis, but play an important role in
systems engineering.

@@

  % Additions from CAV:
  % continuous time
  % floating point 
\item We automatically generate {\em correct-by-construction} digital
  controllers using an inductive synthesis approach.  Our application of program
  synthesis is non-trivial and addresses challenges specific to control
  systems, such as quantizers and FWL.  %% In particular, we
  %% have found that a two-stage verification engine that continuously refines
  %% the precision of the fixed-point representation of the plant yields a
  %% speed-up of two orders of magnitude over a conventional one-stage
  %% verification engine.
%
  %\item  We compute
  The computed state-feedback controllers 
  guarantee a given safety property alongside stability and safety
  specifications. Existing methods for controller synthesis rely on transfer
  function representations, which are inadequate to prove safety requirements.
%
\item We provide two novel algorithms: the first, na\"ive one, relies on
  an unfolding of the dynamics up to a completeness threshold, while the
  second one is abstraction-based and leverages abstraction refinement and
  acceleration to improve scalability while retaining soundness.  %% A third
  %% algorithm evaluates systems using transfer functions where no state-space
  %% description is given.
  Both approaches provide sound synthesis of
  feedback controllers and consider the various sources of imprecision in
  the implementation of the control algorithm and in the modeling of the
  plant.
%
\item We develop a model for different sources of quantization error and
  their effect on reachability properties.  We give bounds that ensure the
  safety of our controllers in a hybrid continuous-digital domain.
\end{enumerate}

@@

%-------------------------------
%\subsection{Formal specification of transient response} 
%\label{ssec:transientspecification}
%-------------------------------

%In a SISO system, there are 2 values that typically define the response-time of the system (using a step response).
%\begin{enumerate}
%\item The \emph{rise time} is defined as the time it takes the output to reach a percentage (eg $90\%$) of the target value given a step chnage in the input.
%\item The \emph{settling time} is defined as the time it takes the output to stabilize within a margin of the target value (eg $10\%$).
%\end{enumerate} 
%In the case of first and second order systems (with ), the decay is governed by the equations 
%\begin{align*}
%&y_{decay}=
%&\left\{
%\begin{array}{cc}
%e^{-|\omega| t}& 1^{st} order\\
%e^{\zeta |\omega| t}\left(\frac{\sin\left(|\omega| t\sqrt{1-\zeta^2}+\sin^{-1}\sqrt{1-\zeta^2}\right)}{\sqrt{1-\zeta^2}} \right)&\zeta<1\\
%e^{-|\omega| t}-|\omega| t e^{-|\omega| t}&\zeta=1\\
%\frac{|\omega| e^{-|\omega_2| t}-|\omega_2| e^{-|\omega| t}}{|\omega|-|\omega_2|}&\zeta>1\end{array}\right.
%\end{align*}
%where $|\omega|$ and $|\omega_2|$ are the magnitudes of the eigenvalues (roots) of the system and  $\zeta=-\cos(\theta)$ where $\theta$ is the polar angle of the conjugate pair if it exists ($\zeta \geq 1$ if it doesn't).
%These equations become increasingly complicated for higher order systems. However, we note their strong dependency on the magnitude of the eigenvalues, and use this property to define an upperbound on the decay function.
%All of the above equations will decay faster than $e^{-|\omega| t}$ given $|\omega| < |\omega_2|$, hence we may use this function as our upper bound.
%\begin{theorem}
%Given a set of eigenvalues for a discrete time model of an $n^{th}$ order system $\lambda_i =e^{\omega_i}: i \in [1\ n]$, the settling time of the system is upperbounded by
%\begin{equation}
%t_s<-log_{|\overline{\lambda}|}({p_s}) : |\overline{\lambda}| =\max(|\lambda_i|) \wedge |\overline{\lambda}|<1,\ \forall i \in [1\ n]
%\label{eq:set_time}
%\end{equation}
%where $0\leq p_s \leq 1$ is the desired percentage of the target within to remain after $t_s$.
%\end{theorem} 
%\begin{proof}
%Given a closed loop single output LTI system of order $n$ with dynamics
%\begin{equation}
%\left\{\begin{array}{c}\vec{x}_k=\mat{A}\vec{x}_{k-1}\\ \vec{y}_k=\mat{C}\vec{x}_k\end{array}\right. \Rightarrow \left\{ \begin{array}{c}\vec{x}_k'=\mat{J}\vec{x}_{k-1}'\\ \vec{y}_k=\mat{C}\mat{S}\vec{x}_k'\end{array}\right. : \left\{ \begin{array}{c}\mat{A}=\mat{S}\mat{J}\mat{S}^{-1} \\ \vec{x}_k'=\mat{S}^{-1}\vec{x}_k\end{array}\right.
%\label{eq:eigendynamics}
%\end{equation}
%
%the output decay is a linear combination of the decay of eigenvalues or
%eigenvalue pairs.  
%Let $\mat{C}'\in \mathbb{R}^{1 \times n}=\mat{C}\mat{S}$. Since $\mat{C}'$ is constant, the output $y_k$ is a linear combination of the states $\vec{x}'_k$. We accelerate equation \eqref{eq:eigendynamics} to obtain
%\begin{align}
%y_k&=\mat{C}'\mat{J}^k\vec{x}_0 = \sum_{i=0}^n \left( c_i \lambda_i^k + c_{il}\binom{k}{l} \lambda_i^{k-l}\right)\\
%|y_k|&\leq \sum_{i=0}^n \left|c_i + c_{il}\binom{k}{l} \lambda_i^{-l}\right||\lambda_i^k|
%\end{align}
%where the second  part of the equation corresponds to the upper diagonal terms of the Jordan blocks. We also have $|\lambda_i^k|\leq |\overline{\omega}^k|$,hence
%\begin{align}
%|y_k|&\leq |\overline{\omega}^k|\sum_{i=0}^n \left|c_i + c_{il}\binom{k}{l} \lambda_i^{-l}\right|
%\end{align}
%Since we are synthesising the eigenvalues, rather than solve this formula for complex Jordan Shapes, we will restrict our synthesiser to provide solutions with distinct eigenvalues (which is by far the most common case). Therefore the right hand terms dissappear and we end up with
%\begin{align}
%|y_k|&\leq |\overline{\omega}^k|\sum_{i=0}^n |c_i|
%\end{align}

%Hence the overall decay of the system is always smaller
%than $p=e^{-|\overline{\omega}| t}$ where $0\leq p\leq 1$ is the remaining
%percentage of the decay expressed.  reversing we have
%$t=-\frac{ln(p)}{|\overline{\omega}|}=-\frac{ln(p)}{ln(|\overline{\lambda}|)}$. 
%replacing for $p_s$ , $t_s$, and $\overline{\lambda}$ we get equation
%\eqref{eq:set_time}
%
%\qed
%\end{proof}
%
%We may now use this limit to express a constraint in terms of the
%coefficients for our synthesis process.  Let
%
%\begin{equation}
%P=\prod_{i=0}^n (s-\lambda_i)=\sum_{j=0}^n c_js^j.
%\end{equation}
%
%We use Jury's criterion to limit $|\lambda_i| \leq 1$.  Let us now define a
%transformation $\lambda_i'=\frac{\lambda_i}{\overline{\lambda}}$.  We want
%to ensure $\lambda_i' \leq 1$ which can be done by applying Jury's criterion
%to $P'=\prod_{i=0}^n (s-\lambda_i')=\sum_{j=0}^n c_j's^j$ where $c_j'$ are
%unknown.  However, if we replace for $\lambda_i$, we obtain
%
%\begin{equation}
%P'=\prod_{i=0}^n (s-\lambda_i')=\prod_{i=0}^n \left(s-\frac{\lambda_i}{\overline{\lambda}}\right)=\sum_{j=0}^n \frac{c_j}{\overline{\lambda}^{n-j}}s^j
%\end{equation}
%
%where all members are known, we may now synthesise coefficients that ensure
%the eigenvalues are smaller than a desired magnitue, thus ensuring that the
%settling time is too.
